---
title: "pre-assignment"
author: "Jakub Skrajny"
date: "27 04 2021"
output:
  pdf_document: default
  html_document: default
---
# Libraries
```{r}
library(openxlsx)
library(tseries)
library(forecast)
library(EnvStats)
library(lmtest)
```

# Read data
```{r cars}
data <- read.xlsx("Lions_Den_data.xlsx")
ts <- ts(unlist(data[2]), start=c(1994, 7), frequency=12)
plot(ts)
```

# Decomposition
```{r}
decompose <- decompose(ts, "multiplicative")
plot(decompose)
```

# Outliers
Rosner's test suggest that there is only one outlier, but on the plot we can see that entire winter 2006 is does not match. 
```{r}
plot(ts, type="o", ylab="coal usage")
rosnerTest(ts, k = 3)
```

# Autocorrelation and stationarity analysis
We can see that time-series is already stationary and is slightly correlated with itself 12 month earlier.
```{r}
#autocorrelation
acf(ts, lag.max = 12)
#stationary test
adf.test(ts)
```

# Prediction
Delete outlier. Lower data from winter 2006.
```{r}
ts_ <- ts
plot(window(ts_, 2003, 2008), type="l", ylab="coal usage")
ts_[137:143] = ts_[137:143] * 0.6
lines(window(ts_, 2003, 2008), type="l", col="blue")
```
Perform new decomposition.
```{r}
decompose_ <- decompose(ts_, "multiplicative")
plot(decompose_)
```
Predict trend.
```{r}
tsTrend <- ts(decompose_$trend, start=c(1994, 7), frequency=12)
fitARIMA <- arima(tsTrend, order=c(1,1,1),seasonal = list(order = c(1,0,0), period = 12),method="ML")
coeftest(fitARIMA) 
yTrend <- predict(fitARIMA,n.ahead = 12)$pred
cTrend <- ts(
  c(tsTrend, yTrend),
  start=start(tsTrend),
  frequency=12
)
plot(cTrend, type="s", ylab="coal usage trend")
```
Extract seasonal.
```{r}
ySeasonal <- window(decompose$seasonal, start=2020, end=c(2020,12))
```
Predict coal usage.
```{r}
Y <- ts(as.numeric(yTrend) * as.numeric(ySeasonal),
        star=start(yTrend), frequency=12)
TS <- ts(c(ts_, Y), start=start(ts_), frequency=12)
plot(TS, type="s", ylab="coal usage",
     main="history and prediction")
lines(Y, type="s", col="blue")
```